---
title: "Bellabeat Data Report"
author: "Ajinkya Desai"
date: "2023-02-01"
output:
  html_document: default
  pdf_document: default
---

##Introduction
The Bellabeat app offers health-related data on activity, sleep, stress, menstrual cycle, and mindfulness habits. By using this app, users can gain a better understanding of their current habits and make healthy choices. The app connects with Bellabeat's smart wellness products, including the Leaf tracker, which can be worn as a bracelet, necklace, or clip and tracks activity, sleep, and stress, and the Time watch, which combines a classic look with smart technology to track activity, sleep, and stress. The Spring water bottle is another product that tracks daily water intake using smart technology and connects to the Bellabeat app.

Bellabeat also provides a membership program that offers personalized guidance on nutrition, activity, sleep, health and beauty, and mindfulness based on the user's lifestyle and goals. The company aims to gain insights into how people use their smart devices and identify growth opportunities by focusing on their products and analyzing FitBit Fitness Tracker Data.

Bellabeat has been investing extensively on digital marketing including Google Search and being active on social media platforms. The co-founder Sršen knows that an analysis of Bellabeat’s available consumer data would reveal more opportunities for growth.

To analyze the data and provide recommendations, I will follow the Data Analysis Process: Ask, Prepare, Process, Analyze, Share, and Act.

###1)Ask:
Questions for the analysis:
1.	What are some current patterns in the utilization of intelligent devices?
2.	In what ways could these patterns be relevant to Bellabeat’s clientele?
3.	How might these patterns contribute to shaping Bellabeat’s marketing approach?
  

###2)Prepare:
Installing and loading of packages:

```{r}
install.packages("tidyverse")
install.packages("here")
install.packages("skimr")
install.packages("janitor")
install.packages("dylyr")
install.packages("ggplot2")

library("tidyverse")
library("here")
library("skimr")
library("janitor")
library("dplyr")
library("ggplot2")
```

Importing datasets:

```{r}
setwd("/cloud/project/Bellabeat Fitness data")
daily_activity <- read_csv("dailyActivity_merged.csv")
daily_calories <- read_csv("dailyCalories_merged.csv")
daily_intensities <- read_csv("dailyIntensities_merged.csv")
daily_steps <- read_csv("dailySteps_merged.csv")
heartrate_sec <- read_csv("heartrate_seconds_merged.csv")
minute_METs <- read_csv("minuteMETsNarrow_merged.csv")
sleep_day <- read_csv("sleepDay_merged.csv")
weight_log <- read_csv("weightLogInfo_merged.csv")
```



###3)Process:
Viewing the data frames:

```{r message=FALSE, warning=FALSE}
head(daily_activity)
head(daily_calories)
head(daily_intensities)
head(daily_steps)
head(heartrate_sec)
head(minute_METs)
head(sleep_day)
head(weight_log)

View(daily_activity)
View(daily_intensities)
View(daily_calories)
View(daily_steps)
```


###4)Analyze:
Summarizing the data:
```{r}
n_distinct(daily_activity$Id)
n_distinct(daily_calories$Id)
n_distinct(daily_intensities$Id)
n_distinct(daily_steps$Id)
n_distinct(heartrate_sec$Id)
n_distinct(minute_METs$Id)
n_distinct(sleep_day$Id)
n_distinct(weight_log$Id)

nrow(daily_activity)
nrow(daily_calories)
nrow(daily_intensities)
nrow(daily_steps)
nrow(heartrate_sec)
nrow(minute_METs)
nrow(sleep_day)
nrow(weight_log)

daily_activity %>%
  select(TotalSteps,TotalDistance,SedentaryMinutes,LightlyActiveMinutes,
         FairlyActiveMinutes,VeryActiveMinutes,Calories) %>%
  summary()

heartrate_sec %>%
  select(Value) %>%
  summary()

minute_METs %>%
  select(METs) %>%
  summary()

sleep_day %>%
  select(TotalTimeInBed, TotalMinutesAsleep,TotalSleepRecords) %>%
  summary()

weight_log %>%
  select(WeightPounds, BMI) %>%
  summary()

```

###6)Share:
```{r}

##The Relationship between Very Active Minutes vs Total Daily Calories Burned
ggplot(data=daily_activity)+
         geom_smooth(mapping=aes(x=VeryActiveMinutes, y=Calories))+
         geom_point(mapping=aes(x=VeryActiveMinutes, y=Calories))+
         labs(title = "The Relationship between Very Active Minutes vs Total Daily Calories Burned", caption = "Data collected by Ajinkya Desai")

##The Relationship between Total Daily Steps vs Total Daily Calories Burned
ggplot(data=daily_activity)+
  geom_smooth(mapping=aes(x=TotalSteps, y=Calories))+
  geom_point(mapping=aes(x=TotalSteps, y=Calories))+
  labs(title = "The Relationship between Total Daily Stepss vs Total Daily Calories Burned", caption = "Data collected by Ajinkya Desai")


##The Relationship between Total  Distance vs Total Daily Calories Burned
ggplot(data=daily_activity)+
  geom_smooth(mapping=aes(x=TotalDistance, y=Calories))+
  geom_point(mapping=aes(x=TotalDistance, y=Calories, color=TotalDistance))+
  labs(title = "The Relationship between Total Distance vs Total Daily Calories Burned", caption = "Data collected by Ajinkya Desai")

##The Relationship between Total Minutes Asleep vs Total Time In Bed 
ggplot(data=sleep_day)+
  geom_smooth(mapping=aes(x=TotalMinutesAsleep, y=TotalTimeInBed))+
  geom_point(mapping=aes(x=TotalMinutesAsleep, y=TotalTimeInBed))+
  labs(title = "The Relationship between Total Minutes Asleep vs Total Time In Bed", caption = "Data collected by Ajinkya Desai")

```


###5)Act:
Bellabeat has achieved success by empowering women through providing data on their activity, sleep, stress, hydration levels, and reproductive health. By analyzing how Fitbit users engage with features, recommendations can be made to enhance Bellabeat’s growth.
To increase its reach, Bellabeat should significantly transform and upgrade its app. Rather than simply offering health data, the app should encourage users to reach their fitness goals and become a social media platform.
The Centers for Disease Control and Prevention (CDC) recommends working out with friends to boost motivation, experiment with workouts, and maintain consistency. The CDC even suggests using social media workout apps to connect with friends and achieve fitness goals. Bellabeat’s app could serve as the social media workout app for women, creating a supportive community of women dedicated to prioritizing their health.

Recommendations for Bellabeat’s app include:
1.	Adding social networking features to enable users to share their favorite workouts, wellness tips, healthy meals, etc.
2.	Enabling users to add friends and view each other’s activity.
3.	Creating weekly fitness and wellness challenges to promote usage.
4.	Allowing health and fitness companies to advertise on the app.
5.	Setting a daily goal of 10,000 steps and sending alert notifications to encourage users to meet the goal.
6.	Recommending users to get at least 7 hours of sleep per night and enabling alert notifications to help them meet the goal.
7.	Suggesting users to engage in 75 minutes of vigorous activity per week and enabling alert notifications to encourage them to meet this goal.
8.	Encouraging users to input their weight and height to track their BMI.
9.	Sending notifications to keep users on track to burn the necessary calories if they express interest in losing weight.
10.	Enabling alert notifications if a user’s resting heart rate deviates significantly from their normal range.
11.	Sending notifications to encourage activity if a user has been awake in bed for an hour.
12.	Sending notifications to encourage activity if a user has been sedentary for an extended period of time.
